---
title: "Forest Sampling"
author: "Karin Wolken"
date: "2017-07-07"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{forestSampling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


# Forest Sampling Package   
  
The Forest Sampling is a package of functions that calculates workups for sampling designs commonly used in Forestry. It is an open source opportunity for collaboration and transparency.  
  
These sampling designs include:
* Simple Random Sampling  
* Cluster Sampling  
* Stratified Sampling  
* Systematic Sampling  
* Two-Stage Sampling  

Variations are included, described in detail [here](insertLinkTo#Functions).  
  
# General Information
  
* All functions require data input. For most functions, this data must be a dataframe formatted in the generalized structure. Select functions can act on vectors (`summarize_simple_random` and `summarize_systematic`).  
* In most cases, additional parameters are necessary. Defaults have been provided to help minimize this.  
  
## Common parameters, detailed:  
* `attribute`: assumes target attribute is called `attr`. Set `attribute` to the column header used for the variable of interest. Not required for vector input.  
* `___Tot`: total number of the ___ unit in the sampling design. Commonly the total number of clusters or plots in the sampling frame.  
* `desiredConfidence`: Percent confidence level (two-sided). Default is set to `0.95`.  

# Functions:  
The following are the basic functions and their associated sampling strategy  
summarize_cluster() -- cluster sampling  
summarize_cluster_discrete() -- cluster sampling with Bernoulli distribution  
summarize_simple_random() -- simple random sampling  
summarize_simple_random_discrete() -- simple random sampling with the Bernoulli distribution  
summarize_stratified() -- stratified sampling  
summarize_systematic() -- systematic sampling  
summarize_two_stage() -- two-stage sampling  
  
## Cluster Sample:  
summarize_cluster() can take in plot data or stand data.  
`plotLevelDataExample <- data.frame(clusterID = c(1, 1, 1, 1, 1, 2, 2, 3, 4, 4, 4, 4, 4, 4, 5, 5, 5, 5, 5), attr = c(1000, 1250, 950, 900, 1005, 1000, 1250, 950, 900, 1005, 1000, 1250, 950, 900, 1005, 1000, 1250, 950, 900), isUsed = c(T, T, T, T, T, T, T, T, T, T, T, T, T, T, F, F, F, F, F))`  
> clusterLevelDataExample = data.frame(clusterID = c(1, 2, 3, 4, 5), clusterElements = c(4, 2, 9, 4, 10), sumAttr = c(1000, 1250, 950, 900, 1005), isUsed = c(T, T, F, T, T))

```{r include = FALSE}
plotLevelDataExample <- data.frame(clusterID = c(1, 1, 1, 1, 1, 2, 2, 3, 4, 4, 4, 4, 4, 4,
                                                 5, 5, 5, 5, 5), 
                                   attr = c(1000, 1250, 950, 900, 1005, 1000, 1250, 950,
                                            900, 1005, 1000, 1250, 950, 900, 1005, 1000,
                                            1250, 950, 900), 
                                   isUsed = c(T, T, T, T, T, T, T, T, T, T, T, T, T, T, F,
                                              F, F, F, F))
```

```{r include = FALSE}
clusterLevelDataExample = data.frame(clusterID = c(1, 2, 3, 4, 5), 
                                     clusterElements = c(4, 2, 9, 4, 10), 
                                     sumAttr = c(1000, 1250, 950, 900, 1005), 
                                     isUsed = c(T, T, F, T, T))
```

